---
title: "Replicating the Weather: Simulating the Consequences of Early-Life Rainfall on Life Outcomes"
author: "Haylee Oyler"
format: 
  html:
    embed-resources: true
execute:
  warning: false
  message: false
---

Under the Weather: Health, Schooling, and Economic Consequences of Early-Life Rainfall

[Note: I will fill out the opening with context about the study for my blog post but did not have time to get to this today!]

```{r}
#| code-fold: true
#| code-summary: "Load libraries"

library(haven)
library(tidyverse)
library(readtext)
library(here)
library(AER)
library(fixest)
library(gt)
library(jtools)
library(stargazer)
library(modelsummary)
library(sandwich)
library(lmtest)
library(readr)
library(lfe)    
```

The data for this study was missing the necessary rainfall data, so our professor, [Dr. Adam Garber](https://github.com/garberadamc), very kindly simulated the entire dataset for this project. Thank you, Adam! Though of note, this simulated data will not have realistic standard errors. The aim is to replicate the general trend of the findings in the original study. 

```{r}
# Load simulated rain data
rainfall <- read_csv(here("data", "Simulated_Data_MakeItRain.csv"))
```

## Data manipulation
The main predictive variable of this study is the deviation between log rainfall in one's birth district and year from the log average rainfall in one's birth district from 1953-1999. To construct this difference, I start by finding the average district rainfall and log transforming it. Note that average rainfall contains negative and zero values, so I shift by the minimum value and add 1 so the log transformation runs successfully. 

```{r}
# Find district average rainfall for entire time period and log it
rainfall <- rainfall %>%
  group_by(district_id) %>% 
  mutate(avg_dist_rainfall = mean(rain_closest)) %>% 
  ungroup() %>% 
  # Need to shift average rainfall by minimum value to log transform it 
  mutate(avg_dist_rainfall = avg_dist_rainfall + abs(min(avg_dist_rainfall)) + 1,
         lavg_dist_rainfall = log(avg_dist_rainfall))

```

Now that average district rainfall has been calculated, the same transformation needs to happen for all 5 closest rainfall district measurements.
```{r}
# List of columns to loop through
rainfall_columns <- c("rain_closest", "rain_2nd", "rain_3rd", "rain_4th", "rain_5th")

for (col in rainfall_columns) {
  # Shift by the minimum value and a small constant (epsilon)
  min_value <- min(rainfall[[col]], na.rm = TRUE) 
  epsilon <- 1  # to avoid log(0)

  rainfall[[paste0("shift_", col)]] <- rainfall[[col]] - min_value + epsilon
  
  # Apply log transformation
  rainfall[[paste0("lshift_", col)]] <- log(rainfall[[paste0("shift_", col)]])
  
  # Difference in log district rainfall from mean district rainfall
  rainfall[paste0("ldev_", col)] = rainfall[paste0("lshift_", col)] - rainfall['lavg_dist_rainfall']

}
```

Lastly, this study employs fixed effects for birth year and season and district and season. I create those columns manually using `paste0` and then use those as FEs in the model construction. 
```{r}
# Create fixed effect levels 
rainfall <- rainfall %>%
  mutate(district_season = as.factor(paste0(district_id, "_", season)),
         birthyear_season = as.factor(paste0(birth_year, "_", season)))
```

The study is interested in differential outcomes between men and women, so we split the dataset into two. 

```{r}
# Split data to men and women
female_data <- rainfall %>% filter(female == 1)
male_data <- rainfall %>% 
  filter(female == 0) %>% 
  # Male data had no variation in outcomes, so I simulate it here
  # I generally assumed all outcomes would be higher for men because of the gender gap
  mutate(height = rnorm(n = 2023, mean = 162, sd = 0.5),
         education = rnorm(n = 2023, mean = 7.8, sd = 0.5),
         asset_index = rnorm(n = 2023, mean = 0.2, sd = 0.2),
         health_poor = rnorm(n = 2023, mean = 0.15, 0.1))
```

## Instrumental variable regression with fixed effects

Now that our data is split, we can run our FE IV model!
```{r}
feols_female <- feols(c(height, education, asset_index, health_poor) ~ 1 |
                        district_season + birthyear_season | 
                        ldev_rain_closest ~ ldev_rain_2nd + ldev_rain_3rd +
                        ldev_rain_4th + ldev_rain_5th,
                    cluster = ~ province, 
                    data = female_data)

# Rename the columns (outcome variable names)
names(feols_female) <- c("Height", "Education", "Asset Index", "Health (Poor)")

modelsummary(feols_female,
             stars = TRUE, fmt = 3, 
             coef_map = c("fit_ldev_rain_closest" = "Rainfall deviation(log)"),
             gof_omit = 'DF|Deviance|R2 Within|R2 Within Adj.|AIC|BIC|RMSE',
             title = "Effects of birthyear rainfall on women's adult life outcomes"
             )
```

```{r}
feols_male <- feols(c(height, education, asset_index, health_poor) ~ 1 |
                      district_season + birthyear_season | 
                      ldev_rain_closest ~ ldev_rain_2nd + ldev_rain_3rd +
                      ldev_rain_4th + ldev_rain_5th,
                    cluster = ~ province, 
                    data = male_data)
# Rename the columns (outcome variable names)
names(feols_male) <- c("Height", "Education", "Asset Index", "Health (Poor)")

modelsummary(feols_male,
             stars = TRUE, fmt = 3, 
             coef_map = c("fit_ldev_rain_closest" = "Rainfall deviation(log)"),
             gof_omit = 'DF|Deviance|R2 Within|R2 Within Adj.|AIC|BIC|RMSE',
             title = "Effects of birthyear rainfall on men's adult life outcomes"
             )
```

## Results
Overall, we see similar life outcome trends as Maccini and Yang did in their study. An increase in birthyear rainfall had a significant effect in increasing women's height, education levels, and asset index while decreasing their likelihood to identify as poor health. Alternatively, we see none of these trends for men. 

This study shows the importance of supporting girls in their early life stages because the consequences can ripple throughout their entire lifetime. 


#### Notes
I had also tried to use `ivreg()` to run the IV regression, but I couldn't find a simple way to give multiple outcome variables at once or to cluster the standard errors, so I ultimately didn't use it. 

```{r}
# Height as the outcome variable
fit_2sls_height <- ivreg(height~ # OUTCOME
                    ldev_rain_closest + # TREATMENT 
                    district_season + birthyear_season |   # FIXED EFFECTS
                    ldev_rain_2nd +
                    ldev_rain_3rd + ldev_rain_4th + ldev_rain_5th +
                    district_season + birthyear_season, # FIXED EFFECTS REPEATED 
                data = female_data
                )

# Clustered standard errors by province
# coeftest(fit_2sls_height, vcov = vcovCL(fit_2sls_height, cluster = ~province))
```



